Reconstructing Propositional Proofs in Type Theory

### 2017-10-24 paper/sections/introduction.tex
[ paper ] fixing intro.

Previous studies have reported proof-reconstruction using proof assistants for classical logic where the development is at a mature stage.
Even though we choose a proof assistant for type theory
where proof verification becomes a type-checking task of the proof-term following the Curry-Howard isomorphism. Then, the proof-reconstruction generates such proof-terms by reconstructing the proof replied by the \ATP.
However, some limitations arises by using type theory as we describe in Section~\ref{sec:limitations-type-theory}.

In this paper, we describe how to reconstruct proofs in type theory
replied by \Metis Prover\footnote{v2.3 release 2017102}.
We ported the subset of the \Metis inference rules for the propositional 
fragment using a syntactical treatment of that logic. The \Metis reasoning was formalized in \Agda~\cite{AgdaProp,AgdaMetis}. 
to reconstruct \Metis proofs, we built a proof-reconstruction tool named
\Athena~\cite{Athena} that is able to generate \Agda\footnote{v2.5.3} proof-terms for \Metis derivations. We have avoided the use of propositions meanings towards a future work to support other formalisms.
% but also to contrast the deep embedding using to write propositions in \Agda 
% with a shallow embedding.

% The proof-reconstruction problem aims to solve this issue.
% One way to address this problem is using another major system in this field, the proof assistants.

% A proof-reconstruction tool using this approach will try to verify the proof delivered by the ATP using the proof assistant.

% In this paper, we purpose one of such proof-reconstruction tools.
% Our approach describes the design and implementation to reconstruct
% proofs of the \Metis prover into a proof assistant for type theory like
% \Agda. We use type theory to guarantee that proofs of the \ATP are valid.

% The main disadvantage of X

% One of the main obstacles to reconstruct proof generated by this tools
% are their

% One of the main obstacles to reconstruct proofs generated by
% automatic provers consists of ...

% However,

% The performance of X is limited by ...

% synopsys

%

% Existing research recognizes the critical role by ...

% What we known about X is largely based on ....

% Few published studied have examined the role of

% focus on a full description of the inference rules in the ATP logic.

% No previous studies has given sufficient considerations to ...

% This paper describes the design and implementation of X...

% This study provides new insights into the automatic reasoning of \Metis prover.
% Understanding the link between the ATP and ITP will help xxxx

% DECIR QUE ES NOVEDOSO EN AGDA Y Q NO ESTA HECHO EN TEORIA DE TIPOS
% SYNTATICA VS SEMANTICA.
% APPLICACIONES

% We describe how to reconstruct proofs delivered by an automatic theorem prover
% (henceforth \ATP)
% and an interactive theorem prover (henceforth
% \ITP).

% One way to reconstruct proofs of \ATPs consists of using an \ITP.
% An \ITP system provide us the logic framework to check and validate
% the reply of the \ATPs since

% A proof-reconstruction tool provides such an integration, translating
% the reply generated by the prover into the formalism of the proof
% assistant.

% Then, reconstructing a proof involves an in-depth
% understanding of the algorithms in the \ATP and the logic specification
% in the \ITP. To begin with the proof-reconstruction, it is
% necessary to have a proof from the \ATP in a consistent
% format, that is, a full script describing step-by-step
% with exhaustive details and without ambiguities, the derivation to
% get the actual proof.

% For problems in classical propositional logic (henceforth \CPL),
% from a list of at least forty\footnote{\ATPs available from the web
% service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just a few
% provers were able to deliver proofs (e.g.,
% \name{CVC4}~\cite{Barrett2011}, \name{SPASS}, and
% \name{Waldmeister}~\cite{hillenbrand1997}) and fewer replied with
% a proof
% in a file format like \TSTP~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}
% (e.g., \name{E}, \Metis, \name{Vampire}, and \name{Z3}),
% \len{LFSC}~\cite{Stump2008}, or the \len{SMT-LIB}~\cite{Bohme2011} format.

%  These
% programs are relevant not only because it helps to spread their
% usage but they also increase the confidence of their users about
% their algorithms and their correctness (see, for example, bugs in
% \ATPs~\cite{Keller2013,Bohme2011,Fleury2014} and
% \cite{Kanso2012}).
% We mention some of these tools in Section~\ref{sec:related-work}.

In this paper,

The formalization of these rules allow us to justify step-by-step \Metis derivations. We implemented this formalization in \Agda and we built
a translator tool, \Athena, to generate automatically the proof-terms
from \Metis proofs.

This paper has been organized in the following way.
We mention in Section~\ref{sec:limitations-type-theory}
some limitations of the proof-reconstruction from the type theory
point of view. In Section~\ref{sec:metis-language-and-proofs}, we introduce the
\Metis prover and its deduction system.
In Section~\ref{sec:proof-reconstruction}, we show our
approach to reconstruct proofs deliver by \Metis.
Related work is described in Section~\ref{sec:related-work}.
Section~\ref{sec:conclusions} summarizes the main findings and suggest future work.

>>>>

Previous studies have reported proof-reconstruction using proof assistants for classical logic where the development is at a mature stage.
Even though we choose a proof assistant for type theory
where proof verification becomes a type-checking task of the proof-term following the Curry-Howard isomorphism. Therefore the proof-reconstruction tool generates proof-terms by reconstructing the proof using the formalization 
the prover in type theory. However, this approach suffers from some limitations of type theory as we describe in Section~\ref{sec:limitations-type-theory}.

In this paper, we describe how to reconstruct proofs in type theory
replied by \Metis prover\footnote{v2.3 release 2017102}.
We ported the subset of the \Metis inference rules for the propositional 
fragment using a syntactical treatment of the logic. The \Metis reasoning was formalized in \Agda in~\cite{AgdaProp,AgdaMetis}. We built a proof-reconstruction tool named \Athena~\cite{Athena} that is able to generate \Agda\footnote{v2.5.3} proof-terms for \Metis derivations. We have avoided the use of propositions meanings towards a future work to support other logics
but also contrast the usage of a deep embedding of the logic with
a shallow embedding.

This paper has been organized in the following way.
In Section~\ref{sec:limitations-type-theory} some limitations of type theory
are discussed for proof-reconstruction point view.
In Section~\ref{sec:metis-language-and-proofs}, we introduce the
\Metis prover and its deduction system.
In Section~\ref{sec:proof-reconstruction}, we show our
approach to reconstruct proofs deliver by \Metis.
Related work is described in Section~\ref{sec:related-work}.
Main findings and suggestion for future work is presented Section~\ref{sec:conclusions}.
